Developmental plasticity to high light intensity amplifies the advantage of amphistomous leaves

### 2025-07-22 ms/ms.qmd
updated pai-aa figure

```{r, aa_cap, echo = FALSE, warning=FALSE, message=FALSE} 
#| label: fig-aacap
#| fig-align: center
#| fig-width: 0.1
#| fig-height: 0.1
#| fig-cap: "\\textbf{Amphistomy advantage (\\aax{}) in wild tomatoes is amplified under simulated sunny growth conditions.} \\aax{} ($y$-axis) is the log-response ratio of photosynthesis in an amphistomatous leaf compared to an otherwise identical pseudohypostomatous leaf. In both panels, estimates are shown for plants grown under simulated shade or sun and measured under high light intensity (\\ppfdequals{2000}) and low light intensity (\\ppfdequals{150}). The dashed line at zero indicates no difference in photosynthesis between amphi- and pseudohypostomatous leaves. (\\textbf{A}) The points are the posterior median \\aax{} for each population, with error bars showing 95% confidence intervals. Within each facet, the populations are arranged by \\aax{} estimated in that growth and measurement condition. (\\textbf{B}) The same estimates of \\aax{} in each combination of growth and measurement light intensity ploted against native plant area index (PAI; $x$-axis, log-scale). The confidence intervals were omitted for visual clarity."

>>>>

```{r, aa_cap, echo = FALSE, warning=FALSE, message=FALSE} 
#| label: fig-aacap
#| fig-align: center
#| fig-width: 0.1
#| fig-height: 0.1
#| fig-cap: "\\textbf{Amphistomy advantage (\\aax{}) in greater wild tomatoes from open habitats and is further amplified by developmental plasticity under sunny growth conditions.} \\aax{} ($y$-axis) is the log-response ratio of photosynthesis in an amphistomatous leaf compared to an otherwise identical pseudohypostomatous leaf. In both panels, estimates are shown for plants grown under simulated shade (brown) or sun (orange) and measured under high light intensity (\\ppfdequals{2000}; open circles) and low light intensity (\\ppfdequals{150}; solid circles). The dashed line at zero indicates no difference in photosynthesis between amphi- and pseudohypostomatous leaves. (\\textbf{A}) The points are the posterior median \\aax{} for each population, with error bars showing 95% confidence intervals. Within each facet, the populations are arranged by \\aax{} estimated in that growth and measurement condition. (\\textbf{B}) The same estimates of \\aax{} in each combination of growth and measurement light intensity ploted against native plant area index (PAI; $x$-axis, log-scale). The confidence intervals were omitted for visual clarity. The median relationship between native PAI ($x$-axis; log-scale) and \\aax{} is shown as a solid line with the shaded region between dashed lines showing the 95% confidence ribbon. The * in the upper left corner of each facet indicates the slope of the linear regression is significantly different from zero; n.s. indicates not significant."

### 2025-07-22 ms/ms.tex
updated pai-aa figure

\caption{\label{fig-aacap}\textbf{Amphistomy advantage (\aax{}) in wild tomatoes is amplified under simulated sunny growth conditions.}
\aax{} (\(y\)-axis) is the log-response ratio of photosynthesis in an
amphistomatous leaf compared to an otherwise identical
pseudohypostomatous leaf. In both panels, estimates are shown for plants
grown under simulated shade or sun and measured under high light
intensity (\ppfdequals{2000}) and low light intensity
(\ppfdequals{150}). The dashed line at zero indicates no difference in
photosynthesis between amphi- and pseudohypostomatous leaves.
(\textbf{A}) The points are the posterior median \aax{} for each
population, with error bars showing 95\% confidence intervals. Within
each facet, the populations are arranged by \aax{} estimated in that
growth and measurement condition. (\textbf{B}) The same estimates of
\aax{} in each combination of growth and measurement light intensity
ploted against native plant area index (PAI; \(x\)-axis, log-scale). The
confidence intervals were omitted for visual clarity.}

@@

All models were using a Bayesian model with HMC sampling in the
probabilistic programming language \emph{Stan} (\emph{59}) using the
\emph{R} package \textbf{brms} version 2.22.0 (\emph{60}). We used
\emph{CmdStan} version 2.36.0 and \textbf{cmdstanr} version 0.9.0
(\emph{61}) to interface with \emph{R} version 4.5.0 (\emph{62}). We
sampled the posterior distribution from a single chain with a minimum of
1000 iterations after 1000 warmup iterations. If necessary, we refit
models with more iterations to decrease the Gelman-Rubin convergence
statistic (\(\hat{R}\)) (\emph{63}) to less than
\(`aa_stats[["max_rhat"]]`\) and the bulk effective sample size (ESS) to
greater than \(`aa_stats[["min_ess"]]`\) for all model parameters. We
also increased the the \texttt{adapt\_delta} parameter until there were
fewer than \(`aa_stats[["n_divergent"]]`\) divergent transitions.

@@

\subsection{MARKER - need to update
table}\label{marker---need-to-update-table}

>>>>

\caption{\label{fig-aacap}\textbf{Amphistomy advantage (\aax{}) in greater wild tomatoes from open habitats and is further amplified by developmental plasticity under sunny growth conditions.}
\aax{} (\(y\)-axis) is the log-response ratio of photosynthesis in an
amphistomatous leaf compared to an otherwise identical
pseudohypostomatous leaf. In both panels, estimates are shown for plants
grown under simulated shade (brown) or sun (orange) and measured under
high light intensity (\ppfdequals{2000}; open circles) and low light
intensity (\ppfdequals{150}; solid circles). The dashed line at zero
indicates no difference in photosynthesis between amphi- and
pseudohypostomatous leaves. (\textbf{A}) The points are the posterior
median \aax{} for each population, with error bars showing 95\%
confidence intervals. Within each facet, the populations are arranged by
\aax{} estimated in that growth and measurement condition. (\textbf{B})
The same estimates of \aax{} in each combination of growth and
measurement light intensity ploted against native plant area index (PAI;
\(x\)-axis, log-scale). The confidence intervals were omitted for visual
clarity. The median relationship between native PAI (\(x\)-axis;
log-scale) and \aax{} is shown as a solid line with the shaded region
between dashed lines showing the 95\% confidence ribbon. The * in the
upper left corner of each facet indicates the slope of the linear
regression is significantly different from zero; n.s. indicates not
significant.}

@@

All models were using a Bayesian model with HMC sampling in the
probabilistic programming language \emph{Stan} (\emph{59}) using the
\emph{R} package \textbf{brms} version 2.22.0 (\emph{60}). We used
\emph{CmdStan} version 2.36.0 and \textbf{cmdstanr} version 0.9.0
(\emph{61}) to interface with \emph{R} version 4.5.0 (\emph{62}). We
sampled the posterior distribution from a single chain with a minimum of
1000 iterations after 1000 warmup iterations. If necessary, we refit
models with more iterations to decrease the Gelman-Rubin convergence
statistic (\(\hat{R}\)) (\emph{63}) to less than
\(`aa_stats[["max_rhat"]]`\) and the bulk effective sample size (ESS) to
greater than \(`aa_stats[["min_ess"]]`\) for all model parameters, and
there were fewer than \(`aa_stats[["n_divergent"]]`\) divergent
transitions.

@@
